---
title: "P8105 Homework 2"
author: "Adina Zhang"
date: "September 30, 2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(p8105.datasets)
```

# Problem 1 - NYC Transit Dataset

```{r transit_df, message = FALSE}
# Read and clean NYC Transit dataset
transit_df <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))

transit_df
```

The NYC Transit dataset includes information on the exits and entrances of various subway stations in the NYC transit network. The variables used to describe this information include railway lines, station names, locations, routes served, entries, entrance types, vending machines, and ADA compliance. Thus far, the data cleaning steps include cleaning the variable names, selecting relevant variables to keep in the dataset, and recoding the *entry* variable into a logical type (TRUE vs. FALSE). The dimension of this dataset is `r nrow(transit_df)` observations with `r ncol(transit_df)` variables.  
  
Tidiness depends on what information you are looking to get out of the dataset. The part of the NYC transit dataset I would consider untidy are the route numbers. The variable names of route numbers should instead be used as values, especially if you are are looking to compare individual subway stations. 

* There are `r dim(distinct(transit_df, station_name, line))[1]` distinct stations.  
* There are `r nrow(filter(transit_df, ada == TRUE))` stations that are ADA compliant.  
* The proportion of station entrances/exits without vending that allow entry is `r nrow(filter(transit_df, vending == "NO", entry == TRUE))/nrow(filter(transit_df, vending == "NO"))`

```{r reformat_route}
# Tidy up route variables using gather
transit_tidy_df = 
 gather(transit_df, key = route_number, value = route_name, route1:route11)

transit_tidy_df
```

* There are `r nrow(filter(distinct(transit_tidy_df, station_name, line, route_name), route_name == "A"))` distinct stations that serve the A train.
* Of the stations that serve the A train, `r nrow(filter(distinct(transit_tidy_df, station_name, line, route_name, ada), route_name == "A", ada == TRUE))` are ADA compliant.  

# Problem 2

## Mr. Trashwheel Dataset

```{r mr_trash_df}
# Load and clean Mr. Trash Wheel dataset
mr_trash_df = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", range = cellranger::cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(date)) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))

mr_trash_df
```

In the Mr. Trash Wheel dataset, there are `r nrow(mr_trash_df)` observations and `r ncol(mr_trash_df)` variables. The median number of sports balls collected by the dumpster in 2016 is `r median(filter(mr_trash_df, year == 2016)$sports_balls)`. Since 2014, the Mr. Trash Wheel has collected `r sum(mr_trash_df$volume_cubic_yards)` which is about `r sum(mr_trash_df$weight_tons)/(nrow(distinct(mr_trash_df, year)))`tons per year.

## Precipitation Dataset

```{r precip_df}
# Load and clean Precipitation 2017 Dataset
prcp_2017_df = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>% 
  mutate(total_prcp = total, year = 2017) %>% 
  select(month, total_prcp, year)

# Load and clean Precipitation 2016 Dataset
prcp_2016_df = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2016 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>% 
  mutate(total_prcp = total, year = 2016) %>% 
  select(month, total_prcp, year)

# Combine 2016 and 2017 precipitation datasets
prcp_tidy_df = bind_rows(prcp_2016_df, prcp_2017_df) %>% 
  mutate(month = month.name[month])

prcp_tidy_df
```

In the joined 2016 and 2017 precipitation dataset, there are `r nrow(prcp_tidy_df)` observations and `r ncol(prcp_tidy_df)` variables which include month, total precipitation, and year. The total precipitation in 2017 was `r sum(filter(prcp_tidy_df, year == 2017)$total_prcp)` inches. 

# Problem 3 - BRFSS Dataset

```{r brfss_df}
# Load and clean BRFSS dataset
brfss_df = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(year:locationdesc, response, data_value) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(excellent_very_good = excellent + very_good)

brfss_df
```

